Asymmetric relationships and their effects on coexistence

### 2023-07-14 paper/RefereeResponses1.tex

\noindent ***\emph{We address your questions and comments below.} \\

\noindent C4) My biggest question is related to the removal of ATAs from a given true bivariate distribution.  This is described in SI Section S2 and shown visually for an example in SI Figure S1.

My first comment related to this is that some form of this text (whcih is relatively short---around half a page in the SI) and, importantly, the Figure should be in the main text.  I think this is such a key part of the definitions that are being used for the rest of the paper that readers will want to get an intuition for what the "partial sharp" distributions look like and how they are arrived at. \\

\noindent ***\emph{We have done what the referee requests, here, and we agree with this request. Thanks for
making the request. This request is somewhat in tension with other requests, from other referees, 
to minimize the quantitative 
content of the main text, and to keep the overall length under the journal limit while adding a bunch of
verbal explanations requested by the other referees. We point this tension out here, and refer back to it in
other responses, below. We also point out that we have not only moved the figure back to the main text,
we have also added a panel to the figure in response to the related comment C13, below.} \\
%DAN: Waiting on Jasmin to add the new panel. In other replies, you can refer to this reply to justify when you
%decide to add explanations to the SI instead of to the main text, and when you have to respond to requests
%to minimize the quant. Also fill in the correct value of CX above.

@@

\noindent ***\emph{Thank you! This is exactly our purpose, here, and we are pleased it has been effective.
Relating back to our response to C23, we think this purpose cannot really be convincingly accomplished
without first stating the lottery model, since the result uses that model. } \\

\noindent C26) L126: Do you do this in the appendix? \\

\noindent ***\emph{No. We can see how it would be done, but we have not formalized the mathematics. 
To avoid making claims that we have not fully supported
with rigorous mathematics or other formal methods, we rephrased to say the iid assumption is 
``probably not necessary''. Thanks for spurring us to do this.} \\

\noindent C27) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\

@@

\noindent ***\emph{No, the equation is correct as is. Please see equation 10 and related equations of 
Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
the $\#$ term because that terms involves both of them varying (though independently).} \\
%Jasmin to check and make sure Dan is right , here. Maybe there is a better place to refer them to to support our claim?

\noindent C28) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\

\noindent ***\emph{Yes, it is a good idea to make it completely clear to the reader that these are 
used distinctly. We now do so, though we do it above this point in the text - it is better, in our view,
to do this the first time we use the hard brackets. So thanks for pointing this out.
We do not know what the referee means by ``using curly under brackets'' so have not done anything
in response to that part of this comment.} \\

\noindent C29) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\

@@

\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
work (and potential for error) 
of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
After all, negative numbers are basic, and represent no real impediment to the average ecologist understanding the 
paper. And we could instead spend that time working on the concepts underlying the next paper.} \\
%Jasmin to consider (don't do yet) how hard/error prone this change would be, taking into account the 
%formulas in the SI? Unless this is MUCH easier and MUCH less error prone than Dan's hunch, I suggest 
%we don't do it.

\noindent C30) L316: So why do you show values of $\sigma >6$? And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA. In 1d it simply appears that the variance scales with the density. 1e is quite unconvincing to me. That’s generally fine, but figure 1 seems to be a conceptual figure, so I wonder whether you don’t have a better example? \\

@@

Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
this is mentioned here,
as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we take
seriously the referee's comment that they are are not convinced there are ATAs in Fig. 1d-e, and ...
%Jasmin, let's discuss what to do here. We could work back to what specific result this is in 
%the original paper and extract from p-value info, or we could come up with a better example,
%or something else, not sure, let's discuss. Maybe also augment the figure to make it clearer?
%How about if Jasmin: 1) adds statistical support to caption; 2) adds to the figure some means of
%hopefully drawing people's attention to the right features of this thing.
} \\

\noindent C31) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\

@@

\noindent ***\emph{There is already a line at $y=0$, you can see it, for instance, on panel (e). 
It gets unavoidably overplotted on panels for which one of the plotted quantities is close to 0. The reason
for different $y$-axis extents is so that you can see the variation within each panel with maximum 
clarity. The main point, here, is to compare the relative contributions of different components
to coexistence, so it's important to show within-plot variation (between lines and across 
values of the $x$-axis) clearly. We thank the referee for expressing their own uncertainty, which was
useful to be aware of.} \\

\noindent C32) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case. \\

@@

\noindent ***\emph{For Fig. 2, the referee is right that ATAs are only rarely important, as explained in the text 
(see first paragraph of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
is usually one of the largest-magnitude terms. Our results as to the importance of ATAs for coexistence 
contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
of section 4.1.} \\

\noindent C33) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\

@@

\noindent ***\emph{The referee appears to have misinterpreted the figure. The ATA terms is, in fact, usually
negative for this figure. The colors of lines corresponding to different terms are in the legend, and, in the caption,
it is reviewed which of these corresponds to the ATA term. Since these items are clearly labelled, and since 
the $y=0$ line is also already present, we are unsure how to reduce the chances for other readers to make
such errors. But we're happy to hear specific suggestions.} \\

\noindent \textbf{Comments of referee 3:} \\

\noindent C34) This is a well performed paper pointing out that asymmetric tail associations (ATA) in a correlation between the environment and competition can be an important mechanisms of coexistence previously overlooked. ATAs can give additional mechanisms that either promote or hinder coexistence in a similar fashion and magnitude to other previously recognised mechanisms such as the storage effect. To illustrate these ideas the authors combine results from mathematical models, simulations, and empirical examples, and show that under a certain combination of parameters ATAs can be a significant mechanism of species coexistence. \\

@@

\noindent ***\emph{Many thanks for the positive feedback!} \\

\noindent C35) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\

@@

\noindent ***\emph{Thank you for the positive feedback!} \\

\noindent C36) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations. Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics. I will give some examples where the message can be improved, but I would suggest the authors to do the same effort themselves in other parts of the manuscript. \\

@@

\noindent ***\emph{We have addressed the referee's specific suggestions below. We have also redoubled our
efforts to make the manuscript as accessible as possible.} \\
%We have to redouble our efforts, as it says. Maybe we also want to point out somewhere that this particular
%sub-field is already a very technical one, but is nevertheless considered a mainstream and very important
%area in ecology?

\noindent C37) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences. \\

@@

\noindent ***\emph{response} \\
%Jasmin, let's discuss. I cannot, myself, say, even now, precisely *how* or *why* ATAs promote or hinder coexistence. 
%In my view, the paper is about showing that they CAN promote or hinder coexistence, and to what extent, so that 
%future work can continue to explore when they do and when they don't, and precisely how. So I want to write a response
%to that effect. However, perhaps you have an intuition, which can be simply explained, as to how ATAs
%either promote or hinder coexistence. I know that's maybe a lot to ask, but if we have such a thing we should consider
%putting it in. Ecologists always want a two-sentence, math-less explanation for everything. If quantum mechanics were
%part of ecology they'd want a two-sentence intuitive explanation for it, even if it was meaningless aside making
%the reader feel better about themselves. Sometimes such a thing does not exist, but when it does exist it can,
%at least sometimes, be useful. For the record, I think most intuitive, verbal explanations of storage effects
%are mostly useless, but perhaps we can do better than those.
%
%Dan and Jasmin to discuss further. Jasmin to dig out old paragraph of text. Jasmin to work on this.

\noindent C38) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ...

@@

\noindent ***\emph{Thank you for the positive feedback!} \\

\noindent C39) ...but how ATAs enter in play is not well solved. It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented. \\

@@

\noindent ***\emph{response} \\
%Same comment again, Jasmin can work on it.

\noindent C40) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor. Could the authors briefly indicate how this could occur? \\

@@

\noindent ***\emph{<It cannot. ATAs can only influence storage effects, not the other mechanisms, as explained on 
lines xyz. Is that enough? Elaborate and re-iterate stuff from the paper.>} \\
%Dan to do

\noindent C41) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\

@@

\noindent ***\emph{<This is a good question. We need to find a place where the models are presented to motivate
them, AND a place later, in the Discussion (or a new SI section called ``Additional discussion points'') where
we talk about why ATAs matter for one but not the other. I think this is just because the
effects of extremes are accentuated in the beta one, and so the effects of associations in the extremes are
accentuated. But that's not unreasonable, since many species have environments which fluctuate between
being suitable and unsuitable.>} \\
%Jasmin to consider this while brainstorming the prior comment.

\noindent C42) The model used for the chemostat is in its formulation quite different compared to the lottery model. Although the technical details are left for the supplementary material, can the authors explain more in the detail in the main text how the importance of ATAs can be used for this particular model. \\

@@

%This referee is asking us to explain a lot of things in more detail. We have to be careful coming across as
%though we have written a paper ``not for a mainstearm audience'', but we can decline some of these requests
%at least for the main text) as not possible because of space constraints. This particular request does not 
%make perfect sense because of grammatic problems. But overall the referee seems to want us to explain the
%effect intuitively, whereas all we've really done is demonstrate its importance. If we want to insert an intuitive 
%explanation, that'd be more than we've tried for. We should talk about whether we have or could make
%something like that.
\noindent ***\emph{response} \\
%Jasmin to also lump this in with her thinking about the earlier ref 3 comments. 
%Q: When you look at E and C marginals for the diatom model, do they have most of their mass in the tails, like the 
%beta-fecundidies lottery model, or most of their mass in the middle, like the normal-fecundities lottery model?

\noindent C43) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system? this needs to be shown otherwise the importance of ATAs will be considered for another dynamics that do not correspond to the observed ones. \\

@@

\noindent ***\emph{response} \\
%I think the earlier papers already showed that, at least for the parameters used there? Check. And the other parameters
%are experimentally manipulable, so this should be OK, but is worth discussing the extent to which ATAs were important
%for the original parameters. The it would be worth noting that the strongest ATA contributions come for parameters
%which were not imposed in the original experiment, so it may be worth repeating the experiment but with these other
%parameters.

\noindent \textbf{Comments of the editor:} \\

\noindent C43) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\

@@

\noindent ***\emph{Thank you for pointing out the novelty.} \\

\noindent C44) The manuscript is very theoretical with a highly technical language hard to follow by a broad range of ecologists. My main suggestion is that the authors need to rewrite it in several parts to make it accessible and convincing. The authors have to make an effort to use a more biological/ecological language with reference to real world potential examples/cases/systems. Some mathematical material could be moved to Supplementary material. \\

### 2023-07-17 paper/Paper.Rnw
line 100 (C9)
continued work on responses and related edits to ms

%Intro the paper by intro'ing ATAs
At the core of ecology is the study of relationships between organisms and their environments.
Ecologists may ask, for instance, how local environment relates to 
population growth rate. To study such relationships and to understand how two variables interact, ecologists 
very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
and direction of association between variables often do not tell the whole story 
\citep{Ghosh_2020_AER}.
To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c all have the same 
Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has previously been referred to as 
``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
regression), but correlations are extremely common, and other common methods also reveal only a small portion 
of the information content of the relationship between variables \citep{Nelsen2006}.
ATAs and related ideas had previously been seldom considered in ecology, but recently several 
insights have been gained by studying how features of associations beyond standard measures 
impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
plankton population density time series. To our knowledge, the implications of ATAs for competition and coexistence have
not yet been investigated. Doing so is the overarching goal of this study.

@@

%Goal para
We previously demonstrated that ATAs are common in nature can influence ecological phenomena
\citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}; and our arguments above
indicate that ATAs may influence species coexistence. Thus, we here seek to quantify the consequences 
of ATAs for competition, coexistence, and the maintenance of biodiversity, by means of the 
following specific goals. 
G1) We will augment MCT to show formally how ATAs play a role in coexistence. 
MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATAs contribute to these components. 
We examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
spatial variation can be considered in future work.
G2) We will explore, using the simple lottery model, the relative magnitudes 
of the contributions to coexistence that come
from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
sometimes coexist with ATAs but not without, or vice versa. This is to help 
illuminate whether ATA contributions to coexistence and biodiversity maintenance
are likely to be negligible or meaningfully large. 
G3) We will provide an empirical example of species coexistence which highlights the role ATAs can play and the
potential importance of ATAs relative to other mechanisms of coexistence.
The example is a diatom system in a chemostat. In the Discussion, we consider whether climate change may alter
ATAs of environmental variables and thereby influence competition and coexistence.
Overall, our study presents a new mechanism of species coexistence
and a means of understanding its theoretical and empirical importance.

@@

We now specify how to decompose storage effects into
contributions due to ATAs and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between rare and common species mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of 
tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[EC]} + \Delta_i^{[E \nat C]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. We will refer to this term as 
the \emph{ATA effect on coexistence}. 
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
correlation \emph{per se}, remaining after having removed the effects of ATAs.
Note that brackets, $[\cdot]$, and parentheses, $(\cdot)$, are used distinctly in our notation,
throughout the paper - they are not interchangeable.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
The new symbol should correspondingly be pronounced ``partial sharp''.
The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
(\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
below for additional details.

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
The value of $\ddot{\rho}_i$ is then determined in one of 
four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.  
For comparison, Fig. \ref{fig:pedagog2} also shows, for the example explored by that figure, 
the distribution $(E_i^\shp,C_{i \bs i}^\shp)$
defined in the previous section (following \cite{Ellner_2016}).
%DAN: Jasmin will be adding that panel.

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/theory_fig_col.pdf}
\caption[Visual guide to removing asymmetries of association]{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATA present. 
(b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
marginals through a composition of the inverse of the standard normal cumulative distribution function (CDF), 
$\varphi^{-1}$, and 
each variable's CDF, $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively.  
(c) A newly generated bivariate normal distribution, $(e_i,c_{i \bs i})$, with correlation $\ddot{\rho}=\rho$
selected to match that of the original distribution (see Theory). (d) Two new variables, 
$E_i^{\nat}$ and  $C_{i \bs i}^{\nat}$,
which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
which have symmetric association structure. These variables are obtained by transforming the marginals from (c), 
$e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
Panel (e) show the distribution $(E_i^\shp,C_{i \bs i}^\shp)$, defined in Theory
following \citep{Ellner_2016}, for comparison.
%DAN: jasmin is adding this new panel
For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 
$\log_{10}$ of probability density on all panels. 
\label{fig:pedagog2}}
\end{figure}

>>>>

%Intro the paper by intro'ing ATAs
At the core of ecology is the study of relationships between organisms and their environments.
Ecologists may ask, for instance, how local environment relates to 
population growth rate. To study such relationships and to understand how two variables interact, ecologists 
very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
and direction of association between variables often do not tell the whole story 
\citep{Ghosh_2020_AER}.
To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c all have the same 
Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has previously been referred to as 
``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
regression), but correlations are extremely common, and other common methods also reveal only a small portion 
of the information content of the relationship between variables \citep{Nelsen2006}.
ATAs and related ideas had previously been seldom considered in ecology, but recently several 
insights have been gained by studying how features of associations beyond standard measures 
impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
plankton population density time series; and many other examples have also been previously explored 
\citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
competition and coexistence have not yet been investigated. Doing so is the overarching goal of this study.

@@

%Goal para
We previously demonstrated that ATAs are common in nature and can influence ecological phenomena
\citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}; and our arguments above
indicate that ATAs may influence species coexistence. Thus, we here seek to quantify the consequences 
of ATAs for competition, coexistence, and the maintenance of biodiversity, by means of the 
following specific goals. 
G1) We will augment MCT to show formally how ATAs play a role in coexistence. 
MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATAs contribute to these components. 
We examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
spatial variation can be considered in future work.
G2) We will explore, using the simple lottery model, the relative magnitudes 
of the contributions to coexistence that come
from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
sometimes coexist with ATAs but not without, or vice versa. This is to help 
illuminate whether ATA contributions to coexistence and biodiversity maintenance
are likely to be negligible or meaningfully large. 
G3) We will provide an empirical example of species coexistence which highlights the role ATAs can play and the
potential importance of ATAs relative to other mechanisms of coexistence.
The example is a diatom system in a chemostat. In the Discussion, we consider whether climate change may alter
ATAs of environmental variables and thereby influence competition and coexistence.
Overall, our study presents a new mechanism of species coexistence
and a means of understanding its theoretical and empirical importance.

@@

We now specify how to decompose storage effects into
contributions due to ATAs and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between rare and common species mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of 
tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[EC]} + \Delta_i^{[E \nat C]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. We will refer to this term as 
the \emph{ATA effect on coexistence}. 
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
correlation \emph{per se}, remaining after having removed the effects of ATAs.
Note that brackets, $[\cdot]$, and parentheses, $(\cdot)$, are used distinctly in our notation,
throughout the paper - they are not interchangeable.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
The new symbol should correspondingly be pronounced ``partial sharp''.
See SI section \ref{SI-SIsect:techcomm} for technical comments on the new notation. 
The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
(\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
below for additional details.

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
The value of $\ddot{\rho}_i$ is then determined in one of 
four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The question of the uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
is addressed in SI section \ref{SI-SIsect:uniqueness}.
The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.  
For comparison, Fig. \ref{fig:pedagog2} also shows, for the example explored by that figure, 
the distribution $(E_i^\shp,C_{i \bs i}^\shp)$
defined in the previous section (following \cite{Ellner_2016}).

@@

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/theory_fig_col2.pdf}
\caption[Visual guide to removing asymmetries of association]{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATAs present. 
(b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
marginals through a composition of the inverse of the standard normal cumulative distribution function (CDF), 
$\varphi^{-1}$, and 
each variable's CDF, $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively.  
(c) A newly generated bivariate normal distribution, $(e_i,c_{i \bs i})$, with correlation $\ddot{\rho}=\rho$
selected to match that of the original distribution (see Theory). (d) Two new variables, 
$E_i^{\nat}$ and  $C_{i \bs i}^{\nat}$,
which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
which have symmetric association structure. These are the distributions used to compute ATA effects, 
$\Delta_i^{[EC]}$ (Theory). 
These variables are obtained by transforming the marginals from (c), 
$e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
Panel (e) shows the distribution $(E_i^\shp,C_{i \bs i}^\shp)$, defined in Theory
following \cite{Ellner_2016}, for comparison. These are the distributions used to compute storage effects,
$\Delta_i^{(EC)}$ (Theory).
For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 
$\log_{10}$ of probability density on all panels. 
\label{fig:pedagog2}}
\end{figure}
